---
title: "FI JCI Insight Analyses"
author: "Akshay Bareja"
date: "`r Sys.Date()`"
toc: true
number-sections: true
self-contained: true
format: 
  revealjs: default
  html:
    df-print: paged
    code-fold: true
    code-tools: true
  pdf: 
    code-line-numbers: true
editor: source
---

```{r}
#| label: load-packages
#| include: false

library(tidyverse)
library(readxl)
library(mmrm)
library(emmeans)
library(tern.mmrm)
library(Rfit)
library(limma)
library(ggforce)
library(ggrepel)
library(DT)
library(gt)
library(ggsignif)
library(patchwork)
library(ggprism)
library(performance)
library(gtsummary)
library(openxlsx)
library(extrafont)
font_import()
#fonts()
```

::: callout-note
## This document contains code and results for the full re-analysis of some of the experiments presented in the manuscript "Food insecurity in young rats induces cardiometabolic disease phenotypes in a sex and diet-dependent manner".
:::

## Figure 1

Fit MMRM models of the same type to each of Figure 1B-F and return results as a concatenated data frame of only fdr-corrected statistically significant contrasts (using a Benjamini-Hochberg adjusted p-value threshold of 0.05)

```{r}
#| label: fig1
#| message: false
#| warning: false
#| cache: true

# function that imports data + runs mmrm
# fits both ar1 and ar1h and picks the model with the smaller AIC
# Note: I am controlling for baseline as described in - https://opensource.nibr.com/bamdd/src/02h_mmrm.html#implementation
fig1_fun <- function(path,
         sheet_name
){
  
  df <- read_excel(path, 
                   sheet = sheet_name, 
                   skip = 2) |> 
    janitor::clean_names() |> 
    tidyr::pivot_longer(
      cols = -day,
      names_to = c("group", "id"),
      names_pattern = "(.+)_._.+_(\\d+)",
      values_to = "val"
    ) |> 
    dplyr::mutate(
      group = factor(group),
      day = factor(day)
    ) |> 
    dplyr::filter(!if_all(
    .cols = c(group, id, val),
    .fns = is.na
  )) # remove rows that have NAs in all these columns. Included to handle some "edge" files (e.g. the Supplemental Figure 1 data)
  
  # pull out day 1 data
  df_base <- df |> 
      dplyr::filter(day == 1) |> 
      dplyr::select(id, val) |> 
      dplyr::rename(
        base = val
      )
  
  # join with original df
  df_mmrm <- df |> 
    dplyr::filter(day != 1) |>
    dplyr::inner_join(
      df_base,
      join_by(id)
    ) |> 
    dplyr::mutate(
      id = factor(id),
      day = factor(day,
                   levels = 2:max(as.numeric(levels(unique(df$day))))),
      group = factor(group,
                     levels = c("fi", "al"))
    )
  
  # fit with ar1
  mmrm_ar1_res <- mmrm(
    formula = val ~ group + day + base + day:group + day:base + ar1(day | id),
    data = df_mmrm
  )
  
  ar1_obj <- summary(mmrm_ar1_res)
  
  # fit with ar1h
  mmrm_ar1h_res <- mmrm(
    formula = val ~ group + day + base + day:group + day:base + ar1h(day | id),
    data = df_mmrm
  )
  
  ar1h_obj <- summary(mmrm_ar1h_res)
  
  if(ar1_obj$aic_list$AIC < ar1h_obj$aic_list$AIC){
    
    mmrm_res <- mmrm_ar1_res
    
  } else {
    
    mmrm_res <- mmrm_ar1h_res
    
  } # pick model with lower AIC
  
  if(ar1_obj$aic_list$AIC < ar1h_obj$aic_list$AIC){
    
    cov_type <- "ar1"
    
  } else {
    
    cov_type <- "ar1h"
    
  } 
  
  mmrm_fit_emm <- emmeans(mmrm_res, 
                          ~ group | day, 
                          lmer.df = "kenward-roger", 
                          weights = "proportional")
  mmrm_fit_contrasts <- contrast(mmrm_fit_emm, 
                                 adjust = "none", 
                                 method = "trt.vs.ctrl", 
                                 ref = 1)
  
  final_res <- data.frame(mmrm_fit_contrasts)  |>  
                as_tibble() |> 
                dplyr::mutate(adj.p.val = p.adjust(p.value, "BH")) |>
                dplyr::filter(adj.p.val < 0.05) |> # only keep adjusted p-values < 0.05
                dplyr::mutate(figure = {{ sheet_name }}) |> 
                dplyr::select(figure, contrast, day, p.value, adj.p.val)
  
  final_res$cov_type <- cov_type
  
  return(final_res)
  
}

# test on one sheet
# fig1_fun(
#   path = "data/Master Data final 08032026.xlsx",
#   sheet_name = "Fig 1B"
# )

# Apply to all sheets
fig1_res_df <- purrr::map(
  .x = c("Fig 1B", "Fig 1C", "Fig 1D", "Fig 1E", "Fig 1F", "Fig 1G", "Fig 1H", "Fig 1I"),
  .f = \(x){
    
    fig1_fun("data/Master Data final 08032026.xlsx",
             sheet_name = x)
    
  }
) |> 
  purrr::list_rbind()

# display full results
fig1_res_df |> 
  dplyr::mutate(p.value = round(p.value, 2),
                adj.p.val = round(adj.p.val, 2)) |> 
  DT::datatable(
    caption = "Figure 1 signficant results (full)"
  )

# number of statistically significant results for each comparison
fig1_res_df |> 
  dplyr::count(figure) |> 
  dplyr::rename(num_sig_results = n) |> 
  gt::gt() |> 
  gt::tab_header(
    title = "Number of comparisons with adjusted p-value < 0.05"
  )
```

## Supplemental Figure 1

Apply `fig1_fun` to Supplemental Figure 1C-D.

```{r}
#| label: supp-fig1
#| message: false
#| warning: false
#| cache: true

# Apply to all sheets
supp_fig1_res_df <- purrr::map(
  .x = c("S. Fig 1C", "S. Fig 1E", "S.Fig 1F"), # need to treat S. Fig 1D separately
  .f = \(x){
    
    fig1_fun("data/Master Data final 08032026.xlsx",
             sheet_name = x)
    
  }
) |> 
  purrr::list_rbind()

# Analyze S. Fig 1D separately
supp_fig1d_df <- read_excel("data/Master Data final 08032026.xlsx", 
                   sheet = "S. Fig 1D", 
                   skip = 1) |> 
    janitor::clean_names() |> 
    dplyr::select(day:last_col()) |> 
    tidyr::pivot_longer(
      cols = -day,
      names_to = c("group", "id"),
      names_pattern = "(.+)_._.+_(\\d+)",
      values_to = "val"
    ) |> 
    dplyr::mutate(
      group = factor(group),
      day = factor(day)
    )

# pull out day 1 data
supp_fig1d_df_base <- supp_fig1d_df |> 
  dplyr::filter(day == 1) |> 
  dplyr::select(id, val) |> 
  dplyr::rename(
    base = val
  )

# join with original df
supp_fig1d_df_mmrm <- supp_fig1d_df |> 
  dplyr::filter(day != 1) |>
  dplyr::inner_join(
    supp_fig1d_df_base,
    join_by(id)
  ) |> 
  dplyr::mutate(
    id = factor(id),
    day = factor(day,
                 levels = 2:max(as.numeric(levels(unique(supp_fig1d_df$day))))),
    group = factor(group,
                   levels = c("fi", "al"))
  )

# fit with ar1
supp_fig1d_mmrm_ar1_res <- mmrm(
  formula = val ~ group + day + base + day:group + day:base + ar1(day | id),
  data = supp_fig1d_df_mmrm
)

supp_fig1d_ar1_obj <- summary(supp_fig1d_mmrm_ar1_res)

# fit with ar1h
supp_fig1d_mmrm_ar1h_res <- mmrm(
  formula = val ~ group + day + base + day:group + day:base + ar1h(day | id),
  data = supp_fig1d_df_mmrm
)

supp_fig1d_ar1h_obj <- summary(supp_fig1d_mmrm_ar1h_res)

if(supp_fig1d_ar1_obj$aic_list$AIC < supp_fig1d_ar1h_obj$aic_list$AIC){
  
  supp_fig1d_mmrm_res <- supp_fig1d_mmrm_ar1_res
  
} else {
  
  supp_fig1d_mmrm_res <- supp_fig1d_mmrm_ar1h_res
  
} # pick model with lower AIC

if(supp_fig1d_ar1_obj$aic_list$AIC < supp_fig1d_ar1h_obj$aic_list$AIC){
  
  supp_fig1d_cov_type <- "ar1"
  
} else {
  
  supp_fig1d_cov_type <- "ar1h"
  
}

supp_fig1d_mmrm_fit_emm <- emmeans(supp_fig1d_mmrm_res, 
                        ~ group | day, 
                        lmer.df = "kenward-roger", 
                        weights = "proportional")
supp_fig1d_mmrm_fit_contrasts <- contrast(supp_fig1d_mmrm_fit_emm, 
                               adjust = "none", 
                               method = "trt.vs.ctrl", 
                               ref = 1)

supp_fig1d_final_res <- data.frame(supp_fig1d_mmrm_fit_contrasts)  |>  
  as_tibble() |> 
  dplyr::mutate(adj.p.val = p.adjust(p.value, "BH")) |>
  dplyr::filter(adj.p.val < 0.05) |> # only keep adjusted p-values < 0.05
  dplyr::mutate(figure = "S. Fig 1D") |> 
  dplyr::select(figure, contrast, day, p.value, adj.p.val)

supp_fig1d_final_res$cov_type <- supp_fig1d_cov_type

# Combine with supp_fig1_res_df
supp_fig1_res_df_full <- supp_fig1_res_df |> 
                          dplyr::bind_rows(supp_fig1d_final_res)

# display full results
supp_fig1_res_df_full |> 
  dplyr::mutate(p.value = round(p.value, 2),
                adj.p.val = round(adj.p.val, 2)) |>
  DT::datatable(
    caption = "Supp. Figure 1 significant results (full)"
  )

# number of statistically significant results for each comparison
supp_fig1_res_df_full |> 
  dplyr::count(figure) |> 
  dplyr::rename(num_sig_results = n) |> 
  gt::gt() |> 
  gt::tab_header(
    title = "Number of comparisons with adjusted p-value < 0.05"
  )
```

## Figure 3

```{r}
#| label: fig3
#| message: false
#| warning: false
#| cache: true

## Box-and-violin plots
# function that performs t-test and produces p-value and figure name
fig3_stats_fun <- function(path,
         sheet_name){
  
  df <- read_excel(path, 
             sheet = sheet_name, skip = 2) |> 
    janitor::clean_names() |> 
    dplyr::select(hour:starts_with("x")) |> 
    dplyr::select(-starts_with("x"))
  
  # compute means
  df_means <- df |> 
    dplyr::summarise(
      across(
        .cols = matches("\\d$"),
        .fns = list(mean = ~mean(.x, na.rm = TRUE)),
        .names = "{.fn}_{.col}"
      )
    ) |> 
    tidyr::pivot_longer(
      cols = everything(),
      names_to = c("group", "sex", "diet", "id"),
      names_pattern = "mean_(.+)_(.)_(.+)_(\\d+)",
      values_to = "val"
    )
  
  # stats test
  p_val <- t.test(
    df_means |>
      dplyr::filter(group == "al") |>
      dplyr::pull(val),
    df_means |>
      dplyr::filter(group == "fi") |>
      dplyr::pull(val)
  )$p.value
  
  # return result
  res_df <- tibble::tibble(
    figure = {{ sheet_name }},
    p_value = p_val
  )
  
  return(res_df)
  
}

# test
# fig3_stats_fun(
#   "data/Master Data final 08032026.xlsx",
#   sheet_name = "Fig 3A-B"
# )

# Apply to all sheets
fig3_res_df <- purrr::map(
  .x = c("Fig 3A-B", "Fig 3C-D", "Fig 3H-I", "Fig 3J-K", "Fig 3 L-M", "Fig 3N-O", "Fig 3P-Q", "Fig 3 R-S"), 
  .f = \(x){
    
    fig3_stats_fun("data/Master Data final 08032026.xlsx",
                    sheet_name = x)
    
  }
) |> 
  purrr::list_rbind()

fig3_res_df

# plots
# function for making box/violin plots with stats info
fig3_plots_fun <- function(path,
                               sheet_name){
  
  df <- read_excel(path, 
                   sheet = sheet_name, skip = 2) |> 
    janitor::clean_names() |> 
    dplyr::select(hour:starts_with("x")) |> 
    dplyr::select(-starts_with("x"))
  
  # compute means
  df_means <- df |> 
    dplyr::summarise(
      across(
        .cols = matches("\\d$"),
        .fns = list(mean = ~mean(.x, na.rm = TRUE)),
        .names = "{.fn}_{.col}"
      )
    ) |> 
    tidyr::pivot_longer(
      cols = everything(),
      names_to = c("group", "sex", "diet", "id"),
      names_pattern = "mean_(.+)_(.)_(.+)_(\\d+)",
      values_to = "val"
    )
  
  # get y-axis title
  column_names <- read_excel(path, 
                             sheet = sheet_name, skip = 1) |> 
                      colnames()
  
  y_axis_title <- column_names[1]
  
  # get Male/Female info
  mf_info <- read_excel(path, 
                        sheet = sheet_name, skip = 2) |> 
    janitor::clean_names() |> 
    dplyr::select(hour:starts_with("x")) |> 
    dplyr::select(-starts_with("x")) |> 
    dplyr::select(last_col()) |> 
    dplyr::rename_with(
      .fn = ~str_extract(.x, "(?<=\\w_).(?=_\\w)")
    ) |> 
    colnames()
  mf_info <- str_to_upper(mf_info)
  
  # plot
  df_means |> 
    ggplot(aes(group, val)) +
    geom_violin(width = 0.5,
                linewidth = 1.5) +
    geom_boxplot(width = 0.15, 
                 alpha=0.2,
                 linewidth = 1.5) +
    theme_prism() +
    labs(title = {{ sheet_name }},
         y = y_axis_title,
         x = "") +
    scale_x_discrete(labels = 
                       c(paste0("AL ", mf_info),
                         paste0("FI ", mf_info))
                     ) +
    theme(axis.text = element_text(face = "bold",
                                   size = 20),
          axis.title = element_text(size = 25),
          axis.title.y = element_text(size = 10,
                                      vjust = 5))
  
}

# test
# fig3_plots_fun(
#   "data/Master Data final 08032026.xlsx",
#   sheet_name = "Fig 3A-B"
# )

# make all plots
purrr::map(
  .x = c("Fig 3A-B", "Fig 3C-D", "Fig 3H-I", "Fig 3J-K", "Fig 3 L-M", "Fig 3N-O", "Fig 3P-Q", "Fig 3 R-S"),
  .f = \(x){

    fig3_plots_fun("data/Master Data final 08032026.xlsx",
                    sheet_name = x)

  }
)

# save plots as individual figures
# purrr::map(
#   .x = c("Fig 3A-B", "Fig 3C-D", "Fig 3H-I", "Fig 3J-K", "Fig 3 L-M", "Fig 3N-O", "Fig 3P-Q", "Fig 3 R-S"),
#   .f = \(x){
# 
#     plot <- fig3_plots_fun("data/Master Data final 08032026.xlsx",
#                             sheet_name = x) +
#               labs(title = "")
# 
#     ggsave(paste0("figures/Figure 3/",
#                   "Fig 3",
#                   str_extract(x, "(?<=\\-)."),
#                   ".tiff"),
#            plot,
#            height = 5,
#            width = 3
#            )
# 
#   }
# )

```

### Figure 3 final submission

```{r}
#| label: fig3-finalsub
#| message: false
#| warning: false
#| cache: true

# update plot function to change font
fig3_plots_fun_finalsub <- function(path,
                               sheet_name){
  
  df <- read_excel(path, 
                   sheet = sheet_name, skip = 2) |> 
    janitor::clean_names() |> 
    dplyr::select(hour:starts_with("x")) |> 
    dplyr::select(-starts_with("x"))
  
  # compute means
  df_means <- df |> 
    dplyr::summarise(
      across(
        .cols = matches("\\d$"),
        .fns = list(mean = ~mean(.x, na.rm = TRUE)),
        .names = "{.fn}_{.col}"
      )
    ) |> 
    tidyr::pivot_longer(
      cols = everything(),
      names_to = c("group", "sex", "diet", "id"),
      names_pattern = "mean_(.+)_(.)_(.+)_(\\d+)",
      values_to = "val"
    )
  
  # get y-axis title
  column_names <- read_excel(path, 
                             sheet = sheet_name, skip = 1) |> 
                      colnames()
  
  y_axis_title <- column_names[1]
  
  # get Male/Female info
  mf_info <- read_excel(path, 
                        sheet = sheet_name, skip = 2) |> 
    janitor::clean_names() |> 
    dplyr::select(hour:starts_with("x")) |> 
    dplyr::select(-starts_with("x")) |> 
    dplyr::select(last_col()) |> 
    dplyr::rename_with(
      .fn = ~str_extract(.x, "(?<=\\w_).(?=_\\w)")
    ) |> 
    colnames()
  mf_info <- str_to_upper(mf_info)
  
  # plot
  df_means |> 
    ggplot(aes(group, val)) +
    geom_violin(width = 0.5,
                linewidth = 1.5) +
    geom_boxplot(width = 0.15, 
                 alpha=0.2,
                 linewidth = 1.5) +
    theme_prism(base_size = 8) +
    labs(title = {{ sheet_name }},
         y = y_axis_title,
         x = "") +
    scale_x_discrete(labels = 
                       c(paste0("AL ", mf_info),
                         paste0("FI ", mf_info))
                     ) +
    theme(axis.text = element_text(face = "plain",

                                   family = "Arial"),
          axis.title = element_text(
                                    family = "Arial"),
          axis.title.y = element_text(
                                      vjust = 5,
                                      family = "Arial"))
  
}

# test
fig3_plots_fun_finalsub(
  "data/Master Data final 08032026.xlsx",
  sheet_name = "Fig 3C-D"
)

# save plots as individual figures
# purrr::map(
#   .x = c("Fig 3A-B", "Fig 3C-D"),
#   .f = \(x){
# 
#     plot <- fig3_plots_fun_finalsub("data/Master Data final 08032026.xlsx",
#                             sheet_name = x) +
#               labs(title = "")
# 
#     ggsave(paste0("figures/Figure 3/",
#                   "Fig 3",
#                   str_extract(x, "(?<=\\-)."),
#                   ".tiff"),
#            plot,
#            height = 5,
#            width = 3,
#            dpi = 350
#            )
# 
#   }
# )
```

## Supplemental Figure 3

```{r}
#| label: supp-fig3
#| message: false
#| warning: false
#| cache: true

# Apply to all sheets
supp_fig3_res_df <- purrr::map(
  .x = c("S. Fig 3A-B", "S. Fig 3C-D", "S. Fig 3E-F", "S. Fig 3G-H", "S. Fig 3I-J", "S. Fig 3K-L", "S. Fig 3M-N", "S. Fig 3O-P"), 
  .f = \(x){
    
    fig3_stats_fun("data/Master Data final 08032026.xlsx",
                    sheet_name = x)
    
  }
) |> 
  purrr::list_rbind()

supp_fig3_res_df

# make all plots
purrr::map(
  .x = c("S. Fig 3A-B", "S. Fig 3C-D", "S. Fig 3E-F", "S. Fig 3G-H", "S. Fig 3I-J", "S. Fig 3K-L", "S. Fig 3M-N", "S. Fig 3O-P"), 
  .f = \(x){
    
    fig3_plots_fun("data/Master Data final 08032026.xlsx",
                    sheet_name = x)
    
  }
)

# save plots as individual figures
# purrr::map(
#   .x = c("S. Fig 3A-B", "S. Fig 3C-D", "S. Fig 3E-F", "S. Fig 3G-H", "S. Fig 3I-J", "S. Fig 3K-L", "S. Fig 3M-N", "S. Fig 3O-P"),
#   .f = \(x){
# 
#     plot <- fig3_plots_fun("data/Master Data final 08032026.xlsx",
#                             sheet_name = x) +
#               labs(title = "")
# 
#     ggsave(paste0("figures/Supp Figure 3/",
#                   "S. Fig 3",
#                   str_extract(x, "(?<=\\-)."),
#                   ".tiff"),
#            plot,
#            height = 5,
#            width = 3)
# 
#   }
# )
```

## Supplemental Figure 4

```{r}
#| label: supp-fig4
#| message: false
#| warning: false
#| cache: true

# test
# fig3_stats_fun(
#   "data/Master Data final 08032026.xlsx",
#   sheet_name = "S. Fig 4 M-N"
# )

# Apply to all sheets
supp_fig4_res_df <- purrr::map(
  .x = c("S. Fig 4A-B", "S. Fig 4C-D", "S. Fig 4G-H", "S. Fig 4I-J", "S. Fig 4K-L", "S. Fig 4 M-N", "S. Figure 4O-P", "S. Figure 4Q-R"), 
  .f = \(x){
    
    fig3_stats_fun("data/Master Data final 08032026.xlsx",
                    sheet_name = x)
    
  }
) |> 
  purrr::list_rbind()

supp_fig4_res_df

# plots

# test
# fig3_plots_fun(
#   "data/Master Data final 08032026.xlsx",
#   sheet_name = "S. Fig 4 M-N"
# )

# make all plots
purrr::map(
  .x = c("S. Fig 4A-B", "S. Fig 4C-D", "S. Fig 4G-H", "S. Fig 4I-J", "S. Fig 4K-L", "S. Fig 4 M-N", "S. Figure 4O-P", "S. Figure 4Q-R"),
  .f = \(x){

    fig3_plots_fun("data/Master Data final 08032026.xlsx",
                    sheet_name = x)

  }
)

# save plots as individual figures
# purrr::map(
#   .x = c("S. Fig 4A-B", "S. Fig 4C-D", "S. Fig 4G-H", "S. Fig 4I-J", "S. Fig 4K-L", "S. Fig 4 M-N", "S. Figure 4O-P", "S. Figure 4Q-R"),
#   .f = \(x){
# 
#     plot <- fig3_plots_fun("data/Master Data final 08032026.xlsx",
#                             sheet_name = x) +
#               labs(title = "")
# 
#     ggsave(paste0("figures/Supp Figure 4/",
#                   "S. Fig 4",
#                   str_extract(x, "(?<=\\-)."),
#                   ".tiff"),
#            plot,
#            height = 5,
#            width = 3)
# 
#   }
# )
```

## Figure 4

Only considering "End of FI" comparisons between AL and FI for Male and Female rats.

The tables below show the results of these comparisons. Also shown are the corresponding volcano plots.

```{r}
#| label: fig4
#| message: false
#| warning: false
#| cache: true

fig_4a <- read_excel("data/Master Data final 08032026.xlsx", 
    sheet = "Fig 4A", skip = 3) |> 
    dplyr::select(Group, `Animal ID`, `TG...14`:`Ghrelin...23`) |> 
    dplyr::rename_with(
      .cols = matches("\\d$"),
      .fn = ~str_extract(.x, "[^\\.]+")
    ) |> 
  tidyr::drop_na()

fig_4b <- read_excel("data/Master Data final 08032026.xlsx", 
    sheet = "Fig 4B", skip = 2) |> 
    dplyr::select(Group, `Animal ID`, `C3...22`:`C20:4...39`) |> 
    dplyr::rename_with(
      .cols = matches("\\d$"),
      .fn = ~str_extract(.x, "[^\\.]+")
    ) |> 
  tidyr::drop_na()

sfig_5a <- read_excel("data/Master Data final 08032026.xlsx", 
    sheet = "S. Figure  5A", skip = 2) |> 
  dplyr::select(Group, `Animal ID`, `Gly...15`:`Arg...25`) |> 
    dplyr::rename_with(
      .cols = matches("\\d$"),
      .fn = ~str_extract(.x, "[^\\.]+")
    ) |> 
  tidyr::drop_na()
  
sfig_5b <- read_excel("data/Master Data final 08032026.xlsx", 
    sheet = "S. Figure 5B", skip = 2) |> 
  dplyr::select(Group, `Animal ID`, `KIV...7`:`KMV...9`) |> 
    dplyr::rename_with(
      .cols = matches("\\d$"),
      .fn = ~str_extract(.x, "[^\\.]+")
    ) |> 
  tidyr::drop_na()
  
sfig_5c <- read_excel("data/Master Data final 08032026.xlsx", 
    sheet = "S. Figure 5C", skip = 2) |> 
  dplyr::select(Group, `Animal ID`, `3-HIB...6`:`2-HB...7`) |> 
    dplyr::rename_with(
      .cols = matches("\\d$"),
      .fn = ~str_extract(.x, "[^\\.]+")
    ) |> 
  tidyr::drop_na()
  
sfig_5d <- read_excel("data/Master Data final 08032026.xlsx", 
    sheet = "S. Figure 5D", skip = 2) |> 
  dplyr::select(Group, `Animal ID`, `C2...5`) |> 
    dplyr::rename_with(
      .cols = matches("\\d$"),
      .fn = ~str_extract(.x, "[^\\.]+")
    ) |> 
  tidyr::drop_na()
  
# combine 
fig_4_endoffidf <- fig_4a |> 
  dplyr::left_join(
    fig_4b,
    join_by(Group, `Animal ID`)
  ) |> 
  dplyr::left_join(
    sfig_5a,
    join_by(Group, `Animal ID`)
  ) |> 
  dplyr::left_join(
    sfig_5b,
    join_by(Group, `Animal ID`)
  ) |> 
  dplyr::left_join(
    sfig_5c,
    join_by(Group, `Animal ID`)
  ) |> 
  dplyr::left_join(
    sfig_5d,
    join_by(Group, `Animal ID`)
  ) 

# reshape, remove metabolites with > 25% 0 values, and for 
# any metabolites with at least one 0 value, add one-half the minimum observed value for that variable JUST to the 0 values
# log2 transform
fig_4_limma <- fig_4_endoffidf |> 
  dplyr::select(-Group) |> 
  tidyr::pivot_longer(
    cols = -`Animal ID`,
    names_to = "metabolite",
    values_to = "abundance"
  ) |> 
  tidyr::pivot_wider(
    names_from = `Animal ID`,
    values_from = abundance
  ) |> # 45 metabolites
  dplyr::distinct() |> # 45 metabolites (checking for duplicates)
  dplyr::rowwise() |>
  dplyr::mutate(prop_zeroes = sum(c_across(-metabolite) == 0)/length(c_across(-metabolite))*100) |>
  dplyr::ungroup() |>
  dplyr::filter(prop_zeroes < 25) |> 
  dplyr::select(-prop_zeroes) |> 
  tidyr::pivot_longer(
    cols = -metabolite,
    names_to = "subject",
    values_to = "abundance"
  ) |> 
  tidyr::pivot_wider(
    names_from = metabolite,
    values_from = abundance
  ) |> 
  dplyr::mutate(across(
    .cols = !c(subject, where(~sum(.x == 0, na.rm = TRUE) == 0)), 
    .fns = ~if_else(.x == 0,
                    .x + 0.5*min(.x[.x > 0], na.rm = TRUE),
                    .x)
    )) |>
  dplyr::mutate(across(!subject, ~log2(.x))) |> # log2 transform
  tidyr::pivot_longer(
    cols = -subject,
    names_to = "metabolite",
    values_to = "log_abundance"
  ) |> 
  tidyr::pivot_wider(
    names_from = subject,
    values_from = log_abundance
  ) 

fig4_limma_mat <- fig_4_limma |> 
                    dplyr::select(-metabolite) |> 
                    as.matrix()
rownames(fig4_limma_mat) <- fig_4_limma$metabolite

# make metadata
fig4_meta <- tibble::tibble(
  
  id = colnames(fig4_limma_mat),
  treatment_sex = c(rep("AL_M", 6), rep("FI_M", 6), rep("AL_F", 6), rep("FI_F", 6))
  
)

# confirm identical order
# identical(colnames(fig4_limma_mat), fig4_meta$id) # TRUE

# make design matrix
fig4_design <- model.matrix(~ 0 + treatment_sex, fig4_meta)

# compute array weights
fig4_w <- arrayWeights(fig4_limma_mat, fig4_design)

# first fit
fig4_fit <- lmFit(fig4_limma_mat, fig4_design, weights = fig4_w)
  
# make contrast  matrix 
fig4_cont.mat <- limma::makeContrasts(
  m_FI_v_AL = treatment_sexFI_M - treatment_sexAL_M,
  f_FI_v_AL = treatment_sexFI_F - treatment_sexAL_F,
  levels = fig4_design
)

# fit contrasts
fig4_fit.cont <- limma::contrasts.fit(
  fig4_fit,
  fig4_cont.mat
)

# robust eBayes step
fig4_efit <- eBayes(fig4_fit.cont, robust = TRUE)

# present results
topTable(fig4_efit, coef = "m_FI_v_AL", number = Inf) |> 
  tibble::rownames_to_column("metabolite") |>
  dplyr::select(metabolite, logFC, P.Value, adj.P.Val) |> 
  dplyr::mutate(
    across(
      .cols = where(is.numeric),
      .fns = ~round(.x, 3)
    )
  ) |> 
  DT::datatable(
    caption = "FI vs AL - Male (End of FI)"
  )

topTable(fig4_efit, coef = "f_FI_v_AL", number = Inf) |> 
  tibble::rownames_to_column("metabolite") |>
  dplyr::select(metabolite, logFC, P.Value, adj.P.Val) |> 
  dplyr::mutate(
    across(
      .cols = where(is.numeric),
      .fns = ~round(.x, 3)
    )
  ) |> 
  DT::datatable(
    caption = "FI vs AL - Female (End of FI)"
  )

# Volcano plots
# NOTE - I am using the magrittr pipe here because the native pipe does not allow me to use `.` within a geom!
set.seed(42) # for geom_text_repel
fig4_m_FI_v_AL <- topTable(fig4_efit, coef = "m_FI_v_AL", number = Inf) %>%
  tibble::rownames_to_column("metabolite") %>%
  dplyr::select(metabolite, logFC, P.Value, adj.P.Val) %>% 
  dplyr::mutate(neg_log10_adjp = -log10(adj.P.Val)) %>% 
  ggplot(aes(logFC, neg_log10_adjp)) +
  geom_point() +
  geom_point(data = . %>% 
               dplyr::filter(logFC < -1 & neg_log10_adjp > 1),
             col = "blue",
             size = 3) +
  geom_point(data = . %>% 
               dplyr::filter(logFC > 1 & neg_log10_adjp > 1),
             col = "red",
             size = 3) +
  geom_text_repel(data = . %>%
                    dplyr::filter(
                      (logFC < -1 & neg_log10_adjp > 1)|
                        (logFC > 1 & neg_log10_adjp > 1)
                    ),
                  aes(label = metabolite),
                  fontface = "bold",
                  size = 5) +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = -log10(0.1), linetype = "dashed") +
  geom_vline(xintercept = 1, linetype = "dotdash") +
  geom_vline(xintercept = -1, linetype = "dotdash") +
  theme_prism() +
  labs(title = "FI vs AL (Male)",
       subtitle = "End of FI",
       y = "-log10(adjusted p-value)")

fig4_f_FI_v_AL <- topTable(fig4_efit, coef = "f_FI_v_AL", number = Inf) %>%
  tibble::rownames_to_column("metabolite") %>%
  dplyr::select(metabolite, logFC, P.Value, adj.P.Val) %>% 
  dplyr::mutate(neg_log10_adjp = -log10(adj.P.Val)) %>%
  ggplot(aes(logFC, neg_log10_adjp)) +
  geom_point() +
  geom_point(data = . %>%
               dplyr::filter(logFC < -1 & neg_log10_adjp > 1),
             col = "blue",
             size = 3) +
  geom_point(data = . %>% 
               dplyr::filter(logFC > 1 & neg_log10_adjp > 1),
             col = "red",
             size = 3) +
  geom_text_repel(data = . %>%
                    dplyr::filter(
                      (logFC < -1 & neg_log10_adjp > 1)|
                        (logFC > 1 & neg_log10_adjp > 1)
                    ),
                  aes(label = metabolite),
                  fontface = "bold",
                  size = 5) +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = -log10(0.1), linetype = "dashed") +
  geom_vline(xintercept = 1, linetype = "dotdash") +
  geom_vline(xintercept = -1, linetype = "dotdash") +
  theme_prism() +
  labs(title = "FI vs AL (Female)",
       subtitle = "End of FI",
       y = "-log10(adjusted p-value)")

fig4_m_FI_v_AL
fig4_f_FI_v_AL

# save plots
# ggsave("figures/Figure 4/fig4_male_FI_vs_AL.tiff",
#        fig4_m_FI_v_AL)
# ggsave("figures/Figure 4/fig4_female_FI_vs_AL.tiff",
#        fig4_f_FI_v_AL)

# save tables
fig4_results <- list(
  "Figure 4B" = topTable(fig4_efit, coef = "m_FI_v_AL", number = Inf) %>%
  tibble::rownames_to_column("metabolite"),
  "Figure 4C" = topTable(fig4_efit, coef = "f_FI_v_AL", number = Inf) %>%
  tibble::rownames_to_column("metabolite")
)

# write.xlsx(fig4_results,
#            "tables/Figure 4/supplementary_table_figure_4.xlsx")
```

### Figure 4 final submission

```{r}
#| label: fig4-finalsub
#| message: false
#| warning: false
#| cache: true

# Updated Volcano plots
# NOTE - I am using the magrittr pipe here because the native pipe does not allow me to use `.` within a geom!
set.seed(42) # for geom_text_repel
fig4_m_FI_v_AL_finalsub <- topTable(fig4_efit, coef = "m_FI_v_AL", number = Inf) %>%
  tibble::rownames_to_column("metabolite") %>%
  dplyr::select(metabolite, logFC, P.Value, adj.P.Val) %>% 
  dplyr::mutate(neg_log10_adjp = -log10(adj.P.Val)) %>% 
  ggplot(aes(logFC, neg_log10_adjp)) +
  geom_point() +
  geom_point(data = . %>% 
               dplyr::filter(logFC < -1 & neg_log10_adjp > 1),
             col = "blue",
             size = 3) +
  geom_point(data = . %>% 
               dplyr::filter(logFC > 1 & neg_log10_adjp > 1),
             col = "red",
             size = 3) +
  geom_text_repel(data = . %>%
                    dplyr::filter(
                      (logFC < -1 & neg_log10_adjp > 1)|
                        (logFC > 1 & neg_log10_adjp > 1)
                    ),
                  aes(label = metabolite),
                  fontface = "plain",
                  size = 5) +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = -log10(0.1), linetype = "dashed") +
  geom_vline(xintercept = 1, linetype = "dotdash") +
  geom_vline(xintercept = -1, linetype = "dotdash") +
  theme_prism(base_size = 8) +
  labs(title = "FI vs AL (Male)",
       subtitle = "End of FI",
       y = "-log10(adjusted p-value)") +
  theme(text = element_text(family = "Arial"),
        axis.title = element_text(size = 14),
        plot.title = element_text(size = 14),
        plot.subtitle = element_text(size = 12))

fig4_f_FI_v_AL_finalsub <- topTable(fig4_efit, coef = "f_FI_v_AL", number = Inf) %>%
  tibble::rownames_to_column("metabolite") %>%
  dplyr::select(metabolite, logFC, P.Value, adj.P.Val) %>% 
  dplyr::mutate(neg_log10_adjp = -log10(adj.P.Val)) %>%
  ggplot(aes(logFC, neg_log10_adjp)) +
  geom_point() +
  geom_point(data = . %>%
               dplyr::filter(logFC < -1 & neg_log10_adjp > 1),
             col = "blue",
             size = 3) +
  geom_point(data = . %>% 
               dplyr::filter(logFC > 1 & neg_log10_adjp > 1),
             col = "red",
             size = 3) +
  geom_text_repel(data = . %>%
                    dplyr::filter(
                      (logFC < -1 & neg_log10_adjp > 1)|
                        (logFC > 1 & neg_log10_adjp > 1)
                    ),
                  aes(label = metabolite),
                  fontface = "plain",
                  size = 5) +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = -log10(0.1), linetype = "dashed") +
  geom_vline(xintercept = 1, linetype = "dotdash") +
  geom_vline(xintercept = -1, linetype = "dotdash") +
  theme_prism(base_size = 8) +
  labs(title = "FI vs AL (Female)",
       subtitle = "End of FI",
       y = "-log10(adjusted p-value)") +
  theme(text = element_text(family = "Arial"),
        axis.title = element_text(size = 14),
        plot.title = element_text(size = 14),
        plot.subtitle = element_text(size = 12))

fig4_m_FI_v_AL_finalsub
fig4_f_FI_v_AL_finalsub

# save plots
# ggsave("figures/Figure 4/fig4_male_FI_vs_AL.tiff",
#        fig4_m_FI_v_AL_finalsub,
#        dpi = 350)
# ggsave("figures/Figure 4/fig4_female_FI_vs_AL.tiff",
#        fig4_f_FI_v_AL_finalsub,
#        dpi = 350)

```

## Figure 5

The tables below show the comparisons for FI vs AL for the male and female subgroups. Also shown are Male vs Female comparisons, averaged across treatment groups.

Also shown are volcano plots for these three comparisons.

```{r}
#| label: fig5
#| message: false
#| warning: false
#| cache: true

# import data
# fig 5
fig_5a <- read_excel("data/Master Data final 08032026.xlsx", 
                      sheet = "Fig 5A", skip = 2) |> 
              dplyr::rename(
                metabolite = 1
              ) |> 
            tidyr::drop_na()

sfig_6b <- read_excel("data/Master Data final 08032026.xlsx",
                      sheet = "S. Figure 6B", skip = 2) |>
              dplyr::rename(
                metabolite = 1
              ) |>
            tidyr::drop_na()

sfig_6c <- read_excel("data/Master Data final 08032026.xlsx",
                      sheet = "S. Figure 6C", skip = 2) |>
              dplyr::rename(
                metabolite = 1
              ) |>
            tidyr::drop_na()

sfig_6d <- read_excel("data/Master Data final 08032026.xlsx",
                      sheet = "S. Figure 6D", skip = 2) |>
              dplyr::rename(
                metabolite = 1
              ) |>
            tidyr::drop_na()

sfig_6e <- read_excel("data/Master Data final 08032026.xlsx",
                      sheet = "S. Figure 6E", skip = 2) |>
              dplyr::rename(
                metabolite = 1
              ) |>
            tidyr::drop_na()

# combine
fig_5 <- dplyr::bind_rows(
            fig_5a,
            sfig_6b,
            sfig_6c,
            sfig_6d,
            sfig_6e
)

# note - there is one duplicate metabolite!
# fig_5 |>
#   dplyr::count(metabolite, sort = TRUE)
# ornithine

# remove duplicate, reshape, remove metabolites with > 25% 0 values, and for 
# any metabolites with at least one 0 value, add one-half the minimum observed value for that variable JUST to the 0 values
# log2 transform
fig_5_limma <- fig_5 |> 
  dplyr::distinct() |> 
  dplyr::rowwise() |>
  dplyr::mutate(prop_zeroes = sum(c_across(-metabolite) == 0)/length(c_across(-metabolite))*100) |>
  dplyr::ungroup() |>
  dplyr::filter(prop_zeroes < 25) |> # this removes two metabolites
  dplyr::select(-prop_zeroes) |> 
  tidyr::pivot_longer(
    cols = -metabolite,
    names_to = "subject",
    values_to = "abundance"
  ) |> 
  tidyr::pivot_wider(
    names_from = metabolite,
    values_from = abundance
  ) |> 
  dplyr::mutate(across(
    .cols = !c(subject, where(~sum(.x == 0, na.rm = TRUE) == 0)), 
    .fns = ~if_else(.x == 0,
                    .x + 0.5*min(.x[.x > 0], na.rm = TRUE),
                    .x)
    )) |>
  dplyr::mutate(across(!subject, ~log2(.x))) |> # log2 transform
  tidyr::pivot_longer(
    cols = -subject,
    names_to = "metabolite",
    values_to = "log_abundance"
  ) |> 
  tidyr::pivot_wider(
    names_from = subject,
    values_from = log_abundance
  ) 

fig5_limma_mat <- fig_5_limma |> 
                    dplyr::select(-metabolite) |> 
                    as.matrix()
rownames(fig5_limma_mat) <- fig_5_limma$metabolite

# make metadata
fig5_meta <- tibble::tibble(
  
  id = colnames(fig5_limma_mat),
  treatment_sex = c(rep("AL_M", 6), rep("FI_M", 6), rep("AL_F", 6), rep("FI_F", 6))
)

# confirm identical order
# identical(colnames(fig5_limma_mat), fig5_meta$id) # TRUE

# make design matrix
fig5_design <- model.matrix(~ 0 + treatment_sex, fig5_meta)

# compute array weights
fig5_w <- arrayWeights(fig5_limma_mat, fig5_design)

# first fit
fig5_fit <- lmFit(fig5_limma_mat, fig5_design, weights = fig5_w)
  
# make contrast  matrix 
fig5_cont.mat <- limma::makeContrasts(
  m_FI_v_AL = treatment_sexFI_M - treatment_sexAL_M,
  f_FI_v_AL = treatment_sexFI_F - treatment_sexAL_F,
  m_v_f = (treatment_sexFI_M + treatment_sexAL_M)/2 - (treatment_sexFI_F + treatment_sexAL_F)/2,
  levels = fig5_design
)

# fit contrasts
fig5_fit.cont <- limma::contrasts.fit(
  fig5_fit,
  fig5_cont.mat
)

# robust eBayes step
fig5_efit <- eBayes(fig5_fit.cont, robust = TRUE)

# present results
topTable(fig5_efit, coef = "m_FI_v_AL", number = Inf) |> 
  tibble::rownames_to_column("metabolite") |>
  dplyr::select(metabolite, logFC, P.Value, adj.P.Val) |> 
  dplyr::mutate(
    across(
      .cols = where(is.numeric),
      .fns = ~round(.x, 3)
    )
  ) |> 
  DT::datatable(
    caption = "FI vs AL - Male"
  )

topTable(fig5_efit, coef = "f_FI_v_AL", number = Inf) |> 
  tibble::rownames_to_column("metabolite") |>
  dplyr::select(metabolite, logFC, P.Value, adj.P.Val) |> 
  dplyr::mutate(
    across(
      .cols = where(is.numeric),
      .fns = ~round(.x, 3)
    )
  ) |> 
  DT::datatable(
    caption = "FI vs AL - Female"
  )

topTable(fig5_efit, coef = "m_v_f", number = Inf) |> 
  tibble::rownames_to_column("metabolite") |>
  dplyr::select(metabolite, logFC, P.Value, adj.P.Val) |> 
  dplyr::mutate(
    across(
      .cols = where(is.numeric),
      .fns = ~round(.x, 3)
    )
  ) |> 
  DT::datatable(
    caption = "Male vs Female"
  )

# Volcano plots
set.seed(42) # for geom_text_repel
fig5_m_FI_v_AL <- topTable(fig5_efit, coef = "m_FI_v_AL", number = Inf) %>% 
  tibble::rownames_to_column("metabolite") %>% 
  dplyr::select(metabolite, logFC, P.Value, adj.P.Val) %>%  
  dplyr::mutate(neg_log10_adjp = -log10(adj.P.Val)) %>%  
  ggplot(aes(logFC, neg_log10_adjp)) +
  geom_point() +
  geom_point(data = . %>%  
               dplyr::filter(logFC < -1 & neg_log10_adjp > 1),
             col = "blue",
             size = 3) +
  geom_point(data = . %>%  
               dplyr::filter(logFC > 1 & neg_log10_adjp > 1),
             col = "red",
             size = 3) +
  geom_text_repel(data = . %>%  
                    dplyr::filter(
                      (logFC < -1 & neg_log10_adjp > 1)|
                        (logFC > 1 & neg_log10_adjp > 1)
                    ),
                  aes(label = metabolite),
                  fontface = "bold",
                  size = 5) +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = -log10(0.1), linetype = "dashed") +
  geom_vline(xintercept = 1, linetype = "dotdash") +
  geom_vline(xintercept = -1, linetype = "dotdash") +
  theme_prism() +
  labs(title = "FI vs AL (Male)",
       y = "-log10(adjusted p-value)")

fig5_f_FI_v_AL <- topTable(fig5_efit, coef = "f_FI_v_AL", number = Inf) %>% 
  tibble::rownames_to_column("metabolite") %>% 
  dplyr::select(metabolite, logFC, P.Value, adj.P.Val) %>%  
  dplyr::mutate(neg_log10_adjp = -log10(adj.P.Val)) %>%  
  ggplot(aes(logFC, neg_log10_adjp)) +
  geom_point() +
  geom_point(data = . %>%  
               dplyr::filter(logFC < -1 & neg_log10_adjp > 1),
             col = "blue",
             size = 3) +
  geom_point(data = . %>%  
               dplyr::filter(logFC > 1 & neg_log10_adjp > 1),
             col = "red",
             size = 3) +
  geom_text_repel(data = . %>%  
                    dplyr::filter(
                      (logFC < -1 & neg_log10_adjp > 1)|
                        (logFC > 1 & neg_log10_adjp > 1)
                    ),
                  aes(label = metabolite),
                  fontface = "bold",
                  size = 5) +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = -log10(0.1), linetype = "dashed") +
  geom_vline(xintercept = 1, linetype = "dotdash") +
  geom_vline(xintercept = -1, linetype = "dotdash") +
  theme_prism() +
  labs(title = "FI vs AL (Female)",
       y = "-log10(adjusted p-value)")

fig5_m_v_f <-topTable(fig5_efit, coef = "m_v_f", number = Inf) %>%  
  tibble::rownames_to_column("metabolite") %>% 
  dplyr::select(metabolite, logFC, P.Value, adj.P.Val) %>%  
  dplyr::mutate(neg_log10_adjp = -log10(adj.P.Val)) %>%  
  ggplot(aes(logFC, neg_log10_adjp)) +
  geom_point() +
  geom_point(data = . %>%  
               dplyr::filter(logFC < -1 & neg_log10_adjp > 1),
             col = "blue",
             size = 3) +
  geom_point(data = . %>%  
               dplyr::filter(logFC > 1 & neg_log10_adjp > 1),
             col = "red",
             size = 3) +
  geom_text_repel(data = . %>% 
                    dplyr::filter(
                      (logFC < -1 & neg_log10_adjp > 1)|
                        (logFC > 1 & neg_log10_adjp > 1)
                    ),
                  aes(label = metabolite),
                  fontface = "bold",
                  size = 5) +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = -log10(0.1), linetype = "dashed") +
  geom_vline(xintercept = 1, linetype = "dotdash") +
  geom_vline(xintercept = -1, linetype = "dotdash") +
  theme_prism() +
  labs(title = "Male vs Female",
       subtitle = "Averaged across treatment groups",
       y = "-log10(adjusted p-value)")

fig5_m_FI_v_AL
fig5_f_FI_v_AL
fig5_m_v_f

# save plots
# ggsave("figures/Figure 5/fig5_male_FI_v_AL.tiff",
#        fig5_m_FI_v_AL)
# ggsave("figures/Figure 5/fig5_female_FI_v_AL.tiff",
#        fig5_f_FI_v_AL)
# ggsave("figures/Figure 5/fig5_male_v_female.tiff",
#        fig5_m_v_f)

# save tables
fig5_results <- list(
  "Figure 5B" = topTable(fig5_efit, coef = "m_FI_v_AL", number = Inf) %>% 
  tibble::rownames_to_column("metabolite"),
  "Figure 5C" = topTable(fig5_efit, coef = "f_FI_v_AL", number = Inf) %>% 
  tibble::rownames_to_column("metabolite"),
  "Figure 5D" = topTable(fig5_efit, coef = "m_v_f", number = Inf) %>%  
  tibble::rownames_to_column("metabolite")
)

# write.xlsx(fig5_results,
#            "tables/Figure 5/supplementary_table_figure_5.xlsx")
```

### Figure 5 final submission

```{r}
#| label: fig5-finalsub
#| message: false
#| warning: false
#| cache: true

# Updated Volcano plots
set.seed(42) # for geom_text_repel
fig5_m_FI_v_AL_finalsub <- topTable(fig5_efit, coef = "m_FI_v_AL", number = Inf) %>% 
  tibble::rownames_to_column("metabolite") %>% 
  dplyr::select(metabolite, logFC, P.Value, adj.P.Val) %>%  
  dplyr::mutate(neg_log10_adjp = -log10(adj.P.Val)) %>%  
  ggplot(aes(logFC, neg_log10_adjp)) +
  geom_point() +
  geom_point(data = . %>%  
               dplyr::filter(logFC < -1 & neg_log10_adjp > 1),
             col = "blue",
             size = 3) +
  geom_point(data = . %>%  
               dplyr::filter(logFC > 1 & neg_log10_adjp > 1),
             col = "red",
             size = 3) +
  geom_text_repel(data = . %>%  
                    dplyr::filter(
                      (logFC < -1 & neg_log10_adjp > 1)|
                        (logFC > 1 & neg_log10_adjp > 1)
                    ),
                  aes(label = metabolite),
                  fontface = "plain",
                  size = 5) +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = -log10(0.1), linetype = "dashed") +
  geom_vline(xintercept = 1, linetype = "dotdash") +
  geom_vline(xintercept = -1, linetype = "dotdash") +
  theme_prism(base_size = 8) +
  labs(title = "FI vs AL (Male)",
       y = "-log10(adjusted p-value)") +
  theme(text = element_text(family = "Arial"),
        axis.title = element_text(size = 14),
        plot.title = element_text(size = 14),
        plot.subtitle = element_text(size = 12))

fig5_f_FI_v_AL_finalsub <- topTable(fig5_efit, coef = "f_FI_v_AL", number = Inf) %>% 
  tibble::rownames_to_column("metabolite") %>% 
  dplyr::select(metabolite, logFC, P.Value, adj.P.Val) %>%  
  dplyr::mutate(neg_log10_adjp = -log10(adj.P.Val)) %>%  
  ggplot(aes(logFC, neg_log10_adjp)) +
  geom_point() +
  geom_point(data = . %>%  
               dplyr::filter(logFC < -1 & neg_log10_adjp > 1),
             col = "blue",
             size = 3) +
  geom_point(data = . %>%  
               dplyr::filter(logFC > 1 & neg_log10_adjp > 1),
             col = "red",
             size = 3) +
  geom_text_repel(data = . %>%  
                    dplyr::filter(
                      (logFC < -1 & neg_log10_adjp > 1)|
                        (logFC > 1 & neg_log10_adjp > 1)
                    ),
                  aes(label = metabolite),
                  fontface = "plain",
                  size = 5) +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = -log10(0.1), linetype = "dashed") +
  geom_vline(xintercept = 1, linetype = "dotdash") +
  geom_vline(xintercept = -1, linetype = "dotdash") +
  theme_prism(base_size = 8) +
  labs(title = "FI vs AL (Female)",
       y = "-log10(adjusted p-value)") +
  theme(text = element_text(family = "Arial"),
        axis.title = element_text(size = 14),
        plot.title = element_text(size = 14),
        plot.subtitle = element_text(size = 12))

fig5_m_v_f_finalsub <-topTable(fig5_efit, coef = "m_v_f", number = Inf) %>%  
  tibble::rownames_to_column("metabolite") %>% 
  dplyr::select(metabolite, logFC, P.Value, adj.P.Val) %>%  
  dplyr::mutate(neg_log10_adjp = -log10(adj.P.Val)) %>%  
  ggplot(aes(logFC, neg_log10_adjp)) +
  geom_point() +
  geom_point(data = . %>%  
               dplyr::filter(logFC < -1 & neg_log10_adjp > 1),
             col = "blue",
             size = 3) +
  geom_point(data = . %>%  
               dplyr::filter(logFC > 1 & neg_log10_adjp > 1),
             col = "red",
             size = 3) +
  geom_text_repel(data = . %>% 
                    dplyr::filter(
                      (logFC < -1 & neg_log10_adjp > 1)|
                        (logFC > 1 & neg_log10_adjp > 1)
                    ),
                  aes(label = metabolite),
                  fontface = "plain",
                  size = 5) +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = -log10(0.1), linetype = "dashed") +
  geom_vline(xintercept = 1, linetype = "dotdash") +
  geom_vline(xintercept = -1, linetype = "dotdash") +
  theme_prism(base_size = 8) +
  labs(title = "Male vs Female",
       subtitle = "Averaged across treatment groups",
       y = "-log10(adjusted p-value)") +
  theme(text = element_text(family = "Arial"),
        axis.title = element_text(size = 14),
        plot.title = element_text(size = 14),
        plot.subtitle = element_text(size = 12))

fig5_m_FI_v_AL_finalsub
fig5_f_FI_v_AL_finalsub
fig5_m_v_f_finalsub

# save plots
# ggsave("figures/Figure 5/fig5_male_FI_v_AL.tiff",
#        fig5_m_FI_v_AL_finalsub,
#        dpi = 350)
# ggsave("figures/Figure 5/fig5_female_FI_v_AL.tiff",
#        fig5_f_FI_v_AL_finalsub,
#        dpi = 350)
# ggsave("figures/Figure 5/fig5_male_v_female.tiff",
#        fig5_m_v_f_finalsub,
#        dpi = 350)

```

<!-- ## Resubmission Material -->

<!-- ```{r} -->
<!-- #| label: resub -->
<!-- #| message: false -->
<!-- #| warning: false -->
<!-- #| cache: true -->

<!-- ## Box-and-violin plots -->
<!-- # function that performs t-test and produces p-value and figure name -->
<!-- resub_stats_fun <- function(path, -->
<!--          sheet_name){ -->

<!--   df <- read_excel(path = path,  -->
<!--                  sheet = sheet_name, skip = 1) |>  -->
<!--   janitor::clean_names() |>  -->
<!--   dplyr::select(-ends_with("hr")) -->

<!--   # compute means -->
<!--   df_means <- df |>  -->
<!--     dplyr::summarise( -->
<!--       across( -->
<!--         .cols = matches("\\d$"), -->
<!--         .fns = list(mean = ~mean(.x, na.rm = TRUE)), -->
<!--         .names = "{.fn}_{.col}" -->
<!--       ) -->
<!--     ) |>  -->
<!--     tidyr::pivot_longer( -->
<!--       cols = everything(), -->
<!--       names_to = c("group", "sex", "diet", "id"), -->
<!--       names_pattern = "mean_(.+)_(.)_(.+)_(\\d+)", -->
<!--       values_to = "val" -->
<!--     ) -->

<!--   # stats test -->
<!--   p_val <- t.test( -->
<!--     df_means |> -->
<!--       dplyr::filter(group == "al") |> -->
<!--       dplyr::pull(val), -->
<!--     df_means |> -->
<!--       dplyr::filter(group == "fi") |> -->
<!--       dplyr::pull(val) -->
<!--   )$p.value -->

<!--   # return result -->
<!--   res_df <- tibble::tibble( -->
<!--     figure = {{ sheet_name }}, -->
<!--     p_value = p_val -->
<!--   ) -->

<!--   return(res_df) -->

<!-- } -->

<!-- # test -->
<!-- # resub_stats_fun( -->
<!-- #   "data/HFHS Energy balance data 07312026.xlsx", -->
<!-- #   sheet_name = "Males" -->
<!-- # ) -->

<!-- # Apply to all sheets -->
<!-- resub_res_df <- purrr::map( -->
<!--   .x = c("Males", "Females"),  -->
<!--   .f = \(x){ -->

<!--     resub_stats_fun("data/HFHS Energy balance data 07312026.xlsx", -->
<!--                     sheet_name = x) -->

<!--   } -->
<!-- ) |>  -->
<!--   purrr::list_rbind() -->

<!-- resub_res_df -->

<!-- # plots -->
<!-- # function for making box/violin plots with stats info -->
<!-- resub_plots_fun <- function(path, -->
<!--                                sheet_name){ -->

<!--   df <- read_excel(path = path,  -->
<!--                  sheet = sheet_name, skip = 1) |>  -->
<!--   janitor::clean_names() |>  -->
<!--   dplyr::select(-ends_with("hr")) -->

<!--   # compute means -->
<!--   df_means <- df |>  -->
<!--     dplyr::summarise( -->
<!--       across( -->
<!--         .cols = matches("\\d$"), -->
<!--         .fns = list(mean = ~mean(.x, na.rm = TRUE)), -->
<!--         .names = "{.fn}_{.col}" -->
<!--       ) -->
<!--     ) |>  -->
<!--     tidyr::pivot_longer( -->
<!--       cols = everything(), -->
<!--       names_to = c("group", "sex", "diet", "id"), -->
<!--       names_pattern = "mean_(.+)_(.)_(.+)_(\\d+)", -->
<!--       values_to = "val" -->
<!--     ) -->

<!--   # get y-axis title -->
<!--   column_names <- read_excel(path,  -->
<!--                              sheet = sheet_name, skip = 1) |>  -->
<!--     colnames() -->

<!--   y_axis_title <- column_names[1] -->

<!--   # get Male/Female info -->
<!--   mf_info <- read_excel(path,  -->
<!--                         sheet = sheet_name, skip = 1) |>  -->
<!--     janitor::clean_names() |>  -->
<!--     janitor::clean_names() |>  -->
<!--     dplyr::select(-ends_with("hr")) |>  -->
<!--     dplyr::select(last_col()) |>  -->
<!--     dplyr::rename_with( -->
<!--       .fn = ~str_extract(.x, "(?<=\\w_).(?=_\\w)") -->
<!--     ) |>  -->
<!--     colnames() -->
<!--   mf_info <- str_to_upper(mf_info) -->

<!--   # plot -->
<!--   df_means |>  -->
<!--     ggplot(aes(group, val)) + -->
<!--     geom_violin(width = 0.5, -->
<!--                 linewidth = 1.5) + -->
<!--     geom_boxplot(width = 0.15,  -->
<!--                  alpha=0.2, -->
<!--                  linewidth = 1.5) + -->
<!--     theme_prism() + -->
<!--     labs(title = {{ sheet_name }}, -->
<!--          y = y_axis_title, -->
<!--          x = "") + -->
<!--     scale_x_discrete(labels =  -->
<!--                        c(paste0("AL ", mf_info), -->
<!--                          paste0("FI ", mf_info)) -->
<!--     ) + -->
<!--     theme(axis.text = element_text(face = "bold", -->
<!--                                    size = 20), -->
<!--           axis.title = element_text(size = 25), -->
<!--           axis.title.y = element_text(size = 10, -->
<!--                                       vjust = 5)) -->

<!-- } -->

<!-- # test -->
<!-- # resub_plots_fun( -->
<!-- #   "data/HFHS Energy balance data 07312026.xlsx", -->
<!-- #   sheet_name = "Males" -->
<!-- # ) -->

<!-- # make all plots -->
<!-- purrr::map( -->
<!--   .x = c("Males", "Females"), -->
<!--   .f = \(x){ -->

<!--     resub_plots_fun("data/HFHS Energy balance data 07312026.xlsx", -->
<!--                     sheet_name = x) -->

<!--   } -->
<!-- ) -->

<!-- # save plots as individual figures -->
<!-- # purrr::map( -->
<!-- #   .x = c("Males", "Females"), -->
<!-- #   .f = \(x){ -->
<!-- #  -->
<!-- #     plot <- resub_plots_fun("data/HFHS Energy balance data 07312026.xlsx", -->
<!-- #                             sheet_name = x) + -->
<!-- #               labs(title = "") -->
<!-- #  -->
<!-- #     ggsave(paste0("figures/Resubmission/", -->
<!-- #                   x, -->
<!-- #                   ".tiff"), -->
<!-- #            plot, -->
<!-- #            height = 5, -->
<!-- #            width = 3 -->
<!-- #            ) -->
<!-- #  -->
<!-- #   } -->
<!-- # ) -->
<!-- ``` -->

